Optimize sequencing depth for shotgun metagenomics of pollination system by rarefaction, using a modular profiling pipeline

### 2021-08-25 MResProject/Latex/Draft.tex
CMEECourseWork

% Metagenomic DNA discoveries after sequencing everything (bacteria, parasites, food, gut) inside a bee
\title{Optimize sequencing depth required for reliable profiling of microbiome by rarefaction and multimodel inference, using shotgun metagenomic data from pollination system and a modular profiling pipeline}
\author{Cong Liu}
\date{2021.8}

@@

  \newpage
  Declaration:
  \newline
  All datasets used in this project are provided by Dr. Peter Graystock and the project was conducted under his supervison. 
  I am responsible for the development and conduction of the analyses presented.

  \newpage
  \linenumbers
  \begin{abstract}
    Bee-associated microbiome plays a crucial role in host health and relevant investigations have been boosted by shotgun metagenomics, which provides unique advantages by capturing diversity of multiple taxonomic clades and providing information on function potentiality. 
    However, utilization of shotgun metagenomics is hindered by complexity of data analysis and high cost of sequencing. 
    % However, bioinformatics for metagenomic data analysis is complex, and the results may be compromised by insufficient sequencing depth. 
    Here, an integrated pipeline combining assembly-dependent and -independent methods was introduced for taxonomic and functional profiling of shotgun metagenomic data, and a framework of rarefaction and multimodel inference was constructed for optimizing sequencing depth. 
    Both the pipeline and the framework were used for analysis of real metagenomic datasets ($2 \times 150 bp$ read pairs) from honey bees, bumble bees and flower washes. 
    The integrated pipeline illustrated taxon composition and metabolic potentiality of metagenome, and provided improvement in species identification compared with an assembly-dependent procedure. 
    Sequencing depth can be optimized for species diversity estimation especially when little emphasis is given to rare species, while functional profiling requires deeper sequencing than taxon profiling. 
    % Rarefaction and model fitting was conducted to compute expected species/function diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to optimize sequencing depth, balancing sequencing effort and capability of estimating diversity reliably. 
    % As a result, functional profiling was demanding about sequencing effort. 
    % None of datasets involved was sufficient for providing almost complete coverage of function diversity. 
    % As for species diversity, the average optimal sequencing depth for reliable estimation of Hill number of order 0, 1 and 2 for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; and for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
    % For flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
    These results are helpful for cost-effective utilization of shotgun metagenomics in investigations of pollination system. 
    % The integrated pipeline and the framework for rarefaction analysis can also be applied for researches on other environments. 
  \end{abstract}

@@

  Pollinators such as bees are crucial to maintain global food security and provide stability of natural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview} but their health is under increasing concern \citep{amiri2020using}. 

@@

  Diverse pathogens and environmental factors threaten bee health and are contributing to their declines. 

@@

  A combination of stressors including parasites, pesticide exposure, invasive species, habitat loss and climate change are all contributing to declines of bee populations. \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}.

@@

  Bees are associated with a community of microorganisms influencing host health \citep{engel2016bee,raymann2018role}. 
  Within the gut of bees, stable microbial communities (microbiomes) play crucial roles in food digestion, parasite defence and chemical detoxification. 
  They mediate food digestion processes including 

@@

  They also provide protection against pathogens including \textit{Crithidia}, 

@@

  \citep{cariveau2014variation,maes2016diet}. 
  Besides, bee-associated symbionts can aid in detoxification. 
  They are involved in resistance to both metal and metalloid toxins including cadmium \citep{rothman2019cadmium}, copper \citep{rothman2020direct} and selenate \citep{rothman2019bumble}. 

@@

  Amplicon sequencing, in which a species-specific barcode region is amplified and sequenced, is a powerful method for investigations of microbiome \citep{abdelfattah2018metabarcoding}.  
  It is vastly used in bee microbiome investigations to illustrate the taxonomic diversity of bee-associated bacteria and fungi (\textit{e.g.} \cite{geldert2021dietary,wang2021gut,powell9field,kapheim2021composition}). 
  However, bees visit numerous niches during environment exploration and foraging activities, and can get contact with diverse eDNA signatures, which provide insight into bee ecology by reflecting interactions between bees and other organisms including bacteria, fungi, plants, arthropods and viruses \citep{bovo2018shotgun,ribani2020honey,bovo2020shotgun,matsuzawa2020application}. 

@@

  As a result, functional capacity needs to be inferred based on reference genomes \citep{asshauer2015tax4fun,douglas2018predicting}. 

@@

  As a result, amplicon sequencing-based inference of functional capacity of bee microbiome may not be reliable.

@@

  Shotgun metagenomics provides an alternative for microbiome investigations to overcome drawbacks of amplicon sequencing. 

@@

  Typical goal of metagenomics is to provide taxonomic and functional profile of microbiome, and there is not a golden standard for bioinformatics of metagenome. 

@@

  Probably because of these shortcomings, assembly is skipped in some researches and short reads are directly proceeded for annotation \citep{tringe2005comparative,abubucker2012metabolic,vermote2018amplicon,bovo2018shotgun}, although the accuracy can be compromised due to low information load of short reads \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling}. 

@@

  The sequencing depth is sufficient for reliable estimation of taxon/FGC diversity if and only if the slope of the model is small \citep{hortal2005ed,chao2012coverage,roswell2021conceptual}. 

@@

  In this project, I utilised metagenomes from three environmental types: the gut of a honey bee (\textit{Apis mellifera}), the gut of a common North American bumble bee (\textit{Bombus impatiens}) and the surface of a flower (\textit{Erigeron annuus}). 

@@

  In the integrated pipeline (Figure \ref{Pipeline}), 
  assembly-dependent taxon profiling is conducted after quality filtering and removing host contamination. 
  To address high negative rate of assembly-dependent search, a reference database of genomes from species present in the assembly is constructed and used for filtering non-host reads. 
  Reads not recruited by the reference database are subjected to assembly-free taxonomic search. 
  This integrated pipeline was used to profile metagenomic datasets involved in this project. 
  % It was shown that the construction of reference database and assembly-free search helped identified species not represented by the assembly. 
  Then in order to estimate optimal sequencing depth for reliable representing taxonomic/functional diversity of metagenomes, I simulated different sequencing depth by rarefaction and profiled subsampled datasets using the integrated pipeline, generating inventories of species and FGCs represented by Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}. 
  Diversity of species and KOs was measured by Hill numbers of order 0, 1 and 2. 
  Sequencing depth was measured by number of clean read pairs (150 bp read length). 
  % The higher the order is, the more sensitive to relative abundances Hill number is. 
  The relationship between Hill numbers and sequencing depth was quantified by fitting and averaging asymptotic species accumulation models, and estimation of optimal sequencing depth was given by the point where the slope of rarefaction curve drops to cut-off values. 
  % Sample would be excluded from optimization of sequencing depth if the original dataset was not sufficient for providing an almost complete inventory of species or KOs, which was characterized by small final slope of rarefaction curve for species or KO richness (Hill number of order 0). 
  % As a result, when measuring species diversity by Hill number of order 0, 1 and 2, the average optimal sequencing depth for honey bee samples are 40.33 million, 18.57 million and 17.45 million; for bumble bee samples, the averages are 42.49 million, 40.33 million and 24.77 million; and for flower surface, 0.88 million sequencing depth is not sufficient for providing an almost complete inventory of species, indicating higher sequencing depth is needed. 
  % As for functional diversity, no datasets involved in this project is big enough for providing an almost complete inventory of KOs, indicating the sequencing depth need to be higher than the maximum of each sample type, \textit{i.e.} 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface. 
  % The integrated pipeline and the framework for rarefaction analysis can also be applied for investigations of microbiomes from other environments, and the estimations of optimal sequencing depth provide guidelines for cost-effective utilization of shotgun metagenomics in investigations of pollination system.

@@

  \section{Materials and Methods}
    \subsection{Samples, DNA extraction and sequencing}
    Samples include four honey bees (two from the same hive (hive 13), one from a hive in the same apiary (hive 15) and one caught foraging), three bumble bees from commercially supplied (Biobest) bumblebee colonies (two from the same colony), and one buffer wash of a wild flower (\textit{Erigeron annuus}).  
    \newline
    DNA extraction was performed as in \cite{graystock2020dominant}, followed by library preparation using a template protocol with Illumina Nextera Library Prep kits. Briefly, this involved tagmentation into fragments of 300 bases before eight samples of 10ng were pooled together and sent to Beijing Genomics Institute for further quality control and sequencing using a full lane in the X-ten platform.

    \subsection{Metagenomic profiling using integrated pipeline}
      \subsubsection{Integrated pipeline}
      The integrated pipeline was designed for shotgun metagenomic profiling, \textit{i.e.} assigning reads to taxa or FGCs. 
      It is separated into five modules (Figure \ref{Pipeline}).
      \newline
      In quality control and host removing, raw sequencing data quality is checked using FastQC v.0.11.5 \citep{andrews2010fastqc} and filtered by Trimmomatic v.0.39 \citep{bolger2014trimmomatic}. 
      Then clean reads are mapped to host genome using Bowtie2 v.2.4.2 \citep{langmead2012fast} and non-host reads are extracted by SAMtools v.1.11 \citep{li2009sequence}. 
      \newline
      The non-host reads are subject to the module of assembly-dependent species identification. 
      \textit{De novo} assembly is conducted by SPAdes v.3.15.2 \citep{prjibelski2020using}. 
      Assembled contigs are aligned to NCBI non-redundant (nr) database by DIAMOND v.2.0.7.145 \citep{buchfink2015fast}, and assigned to taxa by MEGAN6 \citep{huson2007megan}. 
      \newline
      Then fragment recruitment is conducted. 
      A reference database is constructed and it comprises reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding genome annotation in general feature format (gff). 
      For each species represented by assembly, its reference genome dataset, if available, is downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
      Then non-host reads are mapped to the reference database by Bowtie2, and unmapped reads are extracted by SAMtools. 
      \newline
      Reads not recruited by the reference database are subjected to assembly-independent species identification. 
      They are aligned to NCBI nr database through DIAMOND and assigned to taxa by MEGAN6.
      \newline
      Finally, functional annotation is conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      It takes coding sequences (CDSs) of genomes in the reference database and reads subject to assembly-independent species identification as input and assigns them to KOs.
      \newline
      The integrated pipeline was used for analyzing metagenomic datasets involved in this study and details in parameter settings of each module are described in Supplementary \ref{ParameterSetting}.

      \subsubsection{Taxon/function quantification and metabolic pathway reconstruction}
      After profiling, identified species and KOs were quantified by calculating relative sequence abundance, \textit{i.e.} proportion of reads assigned to a species/KO in all reads annotated. 
      For species without available reference genomes, their abundances were calculated using reads assigned to them in assembly-independent search. 
      As for taxa with available reference genomes, they may be identified in both assembly-dependent and -independent search due to strain-specific genomic structures that are not present in reference genomes. 
      Their abundances were calculated by summating number of reads that (1) mapped to coding sequences (CDSs) of reference genomes and (2) assigned to them in assembly-independent search. 
      Reads mapped to non-coding regions were not taken into consideration in order to avoid overestimation since the assembly-independent search was based on aligning reads to nr database, which is composed of proteins. 
      As for KO quantification, CDSs with zero-coverage were excluded. 
      Abundances of KOs were calculated by summating number of reads that (1) mapped to CDSs assigned to KOs and (2) assigned to KOs directly. 
      Extraction of CDSs and calculation of their coverage were conducted by BEDtools v.2.30.0 \citep{quinlan2010bedtools}.
      % \newline
      % In order to investigate metabolic potentialities of metagenome, CDSs with non-zero coverage from reference genomes and reads subject to assembly-independent search were functionally annotated using EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      % It searches query sequences  
      % Identified functional categories were quantified by proportion of reads assigned to a category in all reads annotated (relative sequence abundance). 
      % The number of reads assigned to a functional category equals the sum of reads (1) assigned to the category directly and (2) mapped to CDSs assigned to the category.
      \newline
      Metabolic pathways were inferred based on KOs. 
      Reads assigned to plants and arthropods were not included since they were unlikely to represent living organisms. 
      MinPath v.1.6 was used for pathway inference \citep{ye2009parsimony}. 
      It finds a minimal set of KEGG pathways that can explain all KOs provided as input.

@@

      \subsubsection{Simulating different sequencing depth by subsampling and measuring diversity by Hill numbers}
      Species/KO inventories obtained from different sequencing depth was simulated by rarefaction. 
      % \textit{i.e.} randomly taking a proportion of the original dataset without replacement. 
      Since the ratio between numbers of raw and clean reads is dependent on sequencing process and is not influenced by sample type, sequencing depth here refers to number of clean read pairs to exclude variance caused by different proportion of low quality reads in samples of same type. 
      Besides, the expected ratio between host and non-host reads in a metagenomic dataset is dependent on the DNA sample and not impacted significantly by sequencing depth. 
      Thus the proportion of non-host reads in each simulation is expected to be the same with that in the original dataset. 
      \newline
      Based on these considerations, I randomly subsampled non-host dataset of each sample, taking 10\%-100\% of read pairs at interval of 10\% by \textit{reformat.sh} script of BBmap v.38.90 \citep{bushnell2014bbmap}, and profiled subsampled datasets by the integrated pipeline (Figure \ref{Pipeline}). 
      The sequencing depth of each subsampled dataset equals number of subsampled non-host read pairs divided by ratio between non-host and clean read pairs. 
      Thus, each subsampled dataset of non-host reads is corresponded to an imaginary dataset of clean reads, whose proportion of non-host reads is the same with that of the original metagenomic dataset. 
      \newline
      After profiling subsampled datasets, species/KO diversity was measured by Hill numbers of order $q$, defined as Equation \ref{HillNumbers} \citep{hill1973diversity}.
      \begin{equation}
        D^{(q)} = (\sum_{i}(p_i)^q)^{\frac{1}{1-q}}
        \label{HillNumbers}
      \end{equation}
      $p_i$ represents the relative abundance of $i$th species/KO, and $q$ determines sensitivity to relative abundances. 
      When $q = 0$, abundances are not taken into consideration and $D^{(0)}$ equals species/KO richness. 
      When $q = 1$, Hill number is defined as the limit of Equation \ref{HillNumbers} as $q$ tends to 1 (\ref{HillShannon}) and emphasis is given to species/KOs with general abundances.
      \begin{equation}
        D^{(1)} = e^{-\sum_{i}p_ilogp_i}
        \label{HillShannon}
      \end{equation}
      When $q = 2$, high leverage is provided to abundant species/KOs and Hill number equals the inverse of Simpson index (Equation \ref{HillSimpson}).
      \begin{equation}
        D^{(2)} = \frac{1}{\sum_{i}(p_i)^2}
        \label{HillSimpson}
      \end{equation}

@@

      All Hill numbers behave in the following way: if a proportion of categories in an inventory was removed randomly, all Hill numbers would decrease by that proportion \citep{roswell2021conceptual}. 

@@

    Fungal and viruse species were identified. 

@@

    analysing taxonomic and functional diversity, rarefactiona anlysis was conducted. 

@@

      \caption{Rarefaction curves for species (a, b, c) or KO (d, e, f) richness (Hill number of order 0). 
      The horizonal axis represents sequencing depth, and the vertical axis represents richness. 

@@

      Note that the scale of horizonal axis in subfigure c and f is much smaller than that in other subfigures.}

@@

\textit{Bee\_Bimpatiens\_hv3\_1}&2&8.73   &2.20 &6.16 &0.7051  &6.28 &7.06 &0.8088  &9.45 &7.28 &0.8344  &23.89&7.60 &0.8708\\
\textit{Bee\_Bimpatiens\_hv4\_1}&2&8.33   &3.12 &6.28 &0.7540  &8.52 &7.50 &0.9002  &12.30&7.76 &0.9323  &26.68&8.09 &0.9718\\
\textit{Bee\_Bimpatiens\_hv4\_2}&2&7.12   &2.90 &5.34 &0.7493  &7.47 &6.38 &0.8961  &10.58&6.60 &0.9270  &23.73&6.90 &0.9683\\

>>>>

% Metagenomic DNA discoveries after sequencing everything (bacteria, parasites, food, gut) inside a bee
\title{Optimize sequencing depth for shotgun metagenomics of pollination system by rarefaction, using a modular profiling pipeline}
\author{Cong Liu}
\date{2021.8}

@@

  \newpage
  Declaration:
  \newline
  All datasets used in this project are provided by Dr. Peter Graystock and the project was conducted under his supervision. 
  I am responsible for the development and conduction of the analyses presented.

  \newpage
  \linenumbers
  \begin{abstract}
    Bee gut microbiome plays a crucial role in host health, involving in food digestion, pathogen defence and chemical detoxification. 
    Vast relevant investigations use amplicon sequencing to explore bacterial diversity, while emergence of shotgun metagenomics offers unique advantages by capturing diversity of multiple taxonomic clades and providing information on function potentiality. 
    However, utilization of shotgun metagenomics is hindered by complexity of data analysis and high cost of sequencing. 
    % However, bioinformatics for metagenomic data analysis is complex, and the results may be compromised by insufficient sequencing depth. 
    Here, an integrated pipeline combining assembly-dependent and -independent methods was introduced for taxonomic and functional profiling of shotgun metagenomic data, and a framework of rarefaction and multimodel inference was constructed for optimizing sequencing depth. 
    Both the pipeline and the framework were used for analysis of deep-sequencing datasets ($2 \times 150 $bp read pairs) from of honey bees, bumble bees and flower washes. 
    The integrated pipeline illustrated taxon composition and metabolic potentiality of the metagenomes, and provided significant improvements in species identification. 
    For species diversity estimation, about 40 and 43 million read pairs would be sufficient for metagenomic datasets from honey bees and bumble bees, respectively. 
    If low leverage is given to rare species, shallower sequencing can be adopted. 
    Function profiling is much less stable to sequencing depth than species content. 
    Additional functional gene clusters (FGCs) were still being discovered at original sequencing depth of all samples. 
    % Sequencing depth can be optimized for species diversity estimation especially when little emphasis is given to rare species, while functional profiling requires deeper sequencing than taxon profiling. 
    % Rarefaction and model fitting was conducted to compute expected species/function diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to optimize sequencing depth, balancing sequencing effort and capability of estimating diversity reliably. 
    % As a result, functional profiling was demanding about sequencing effort. 
    % None of datasets involved was sufficient for providing almost complete coverage of function diversity. 
    % As for species diversity, the average optimal sequencing depth for reliable estimation of Hill number of order 0, 1 and 2 for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; and for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
    % For flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
    Overall, this project provides guidelines for defining a balance between sequencing cost and acquirement of reliable results for investigations of pollination system. 
    Similar studies for other host species are recommended before undertaking metagenomic projects involving a large number of samples. 
    % The integrated pipeline and the framework for rarefaction analysis can also be applied for researches on other environments. 
  \end{abstract}

@@

  Pollinators such as bees are crucial to maintain global food security and provide stability of natural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview} but their populations are facing declines \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}. 

@@

  % Diverse pathogens and environmental factors threaten bee health and are contributing to their declines. 

@@

  A combination of stressors including parasites, pesticide exposure, invasive species, habitat loss and climate change is contributing to declines of bee populations. \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}.

@@

  % Bees are associated with a community of microorganisms influencing host health \citep{engel2016bee,raymann2018role}. 
  Within the gut of bees, stable microbial communities (microbiomes) play crucial roles in food digestion, parasite defence and chemical detoxification \citep{moran2015genomics,engel2016bee,raymann2018role}. 
  They help mediate food digestion processes including 

@@

  Bee microbiomes also provide protection against pathogens including \textit{Crithidia}, 

@@

  \citep{cariveau2014variation,maes2016diet}, 
  and are involved in resistance to both metal and metalloid toxins including cadmium \citep{rothman2019cadmium}, copper \citep{rothman2020direct} and selenate \citep{rothman2019bumble}. 

@@

  Amplicon sequencing, in which a species-specific barcode region is amplified and sequenced, is a powerful and vastly used method for investigations of microbiome compositions \citep{abdelfattah2018metabarcoding}.  
  In bee microbiome investigations, it is used to illustrate the taxonomic diversity of bee-associated bacteria and fungi (\textit{e.g.} \cite{geldert2021dietary,wang2021gut,powell9field,kapheim2021composition}). 
  However, bees visit numerous niches during environment exploration and foraging activities, and can get contact with diverse eDNA signatures, which provide insight into pollinator ecology by reflecting interactions between bees and other organisms including bacteria, fungi, plants, arthropods and viruses \citep{bovo2018shotgun,ribani2020honey,bovo2020shotgun,matsuzawa2020application}. 

@@

  As a result, functional capacity needs to be inferred based on taxon composition. 
  Amplicon-based function predictors such as Tax4Fun \citep{asshauer2015tax4fun} and PICRUSt \citep{douglas2018predicting} predict functional capacity based on pre-sequenced genomes without taking gene content variation within taxa. 

@@

  As a result, amplicon-based inference of functional capacity of bee microbiome may not be reliable.

@@

  Shotgun metagenomics provides a solution for microbiome investigations to overcome drawbacks of amplicon sequencing. 

@@

  The goal of metagenomics is typically to provide a taxonomic and functional profile of the microbiome, and there is not a gold standard for performing metagenomic data analysis. 
  % bioinformatics of metagenome. 

@@

  Probably because of these shortcomings, assembly is sometimes skipped and short reads are directly proceeded for annotation \citep{tringe2005comparative,abubucker2012metabolic,vermote2018amplicon,bovo2018shotgun}, although the accuracy can be compromised due to low information load of short reads \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling}. 

@@

  The sequencing depth is sufficient for reliable estimation of taxon/FGC diversity if and only if the slope of the rarefaction curve is small \citep{hortal2005ed,chao2012coverage,roswell2021conceptual}. 

@@

  In this project, I utilised metagenomic datasets ($2 \times 150 $bp) from three environmental types: the gut of honey bees (\textit{Apis mellifera}), the gut of common North American bumble bees (\textit{Bombus impatiens}) and the surface of a wild flower (\textit{Erigeron annuus}). 

@@

  The integrated pipeline provided improvement in species identification compared with the assembly-dependent method. 
  Rarefaction analysis showed that about 40/43 million clean read pairs would be a suitable compromise for sequencing honey/bumble bee samples and species diversity detection. 
  Shallower sequencing can be adopted with reduced emphasis on rare species. 
  Functional profiling is more demanding about sequencing depth than species profiling. 
  Significant accumulation of FGCs was observed at final points of rarefaction curves from all samples.  
  These results provide guidelines for defining a balance between sequencing cost and obtaining reliable taxonomic/functional profiles for metagenomic investigations of pollination system. 
  Similar studies for other host species are recommended before undertaking metagenomic projects with big sample size. 
  % In the integrated pipeline (Figure \ref{Pipeline}), 
  % assembly-dependent taxon profiling is conducted after quality filtering and removing host contamination. 
  % To address high negative rate of assembly-dependent search, a reference database of genomes from species present in the assembly is constructed and used for filtering non-host reads. 
  % Reads not recruited by the reference database are subjected to assembly-free taxonomic search. 
  % This integrated pipeline was used to profile metagenomic datasets involved in this project. 
  % % It was shown that the construction of reference database and assembly-free search helped identified species not represented by the assembly. 
  % Then in order to estimate optimal sequencing depth for reliable representing taxonomic/functional diversity of metagenomes, I simulated different sequencing depth by rarefaction and profiled subsampled datasets using the integrated pipeline, generating inventories of species and FGCs represented by Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}. 
  % Diversity of species and KOs was measured by Hill numbers of order 0, 1 and 2. 
  % Sequencing depth was measured by number of clean read pairs (150 bp read length). 
  % % The higher the order is, the more sensitive to relative abundances Hill number is. 
  % The relationship between Hill numbers and sequencing depth was quantified by fitting and averaging asymptotic species accumulation models, and estimation of optimal sequencing depth was given by the point where the slope of rarefaction curve drops to cut-off values. 
  % % Sample would be excluded from optimization of sequencing depth if the original dataset was not sufficient for providing an almost complete inventory of species or KOs, which was characterized by small final slope of rarefaction curve for species or KO richness (Hill number of order 0). 
  % % As a result, when measuring species diversity by Hill number of order 0, 1 and 2, the average optimal sequencing depth for honey bee samples are 40.33 million, 18.57 million and 17.45 million; for bumble bee samples, the averages are 42.49 million, 40.33 million and 24.77 million; and for flower surface, 0.88 million sequencing depth is not sufficient for providing an almost complete inventory of species, indicating higher sequencing depth is needed. 
  % % As for functional diversity, no datasets involved in this project is big enough for providing an almost complete inventory of KOs, indicating the sequencing depth need to be higher than the maximum of each sample type, \textit{i.e.} 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface. 
  % % The integrated pipeline and the framework for rarefaction analysis can also be applied for investigations of microbiomes from other environments, and the estimations of optimal sequencing depth provide guidelines for cost-effective utilization of shotgun metagenomics in investigations of pollination system.

  \section{Materials and Methods}
    \subsection{Samples, DNA extraction and sequencing}
    Samples include four honey bees (two from the same hive (hive 13), one from a hive in the same apiary (hive 15) and one caught foraging), three bumble bees from commercially supplied (Biobest) bumblebee colonies (two from the same colony), and one buffer wash of a wild flower (\textit{Erigeron annuus}).  
    \newline
    DNA extraction was performed as in \cite{graystock2020dominant}, followed by library preparation using a low template protocol with Illumina Nextera Library Prep kits. Briefly, this involved tagmentation into fragments of 300 bases before eight samples of 10ng were pooled together and sent to Beijing Genomics Institute (BGI) for further quality control and sequencing using a full lane in the X-ten platform.

    \subsection{Metagenomic profiling using integrated pipeline}
      \subsubsection{Integrated pipeline}

@@

      I designed an integrated pipeline for shotgun metagenomic profiling, \textit{i.e.} assigning reads to taxa or FGCs. 
      It is separated into five modules (Figure \ref{Pipeline}).
      \newline
      In quality control and host removing, raw sequencing data quality is checked using FastQC v.0.11.5 \citep{andrews2010fastqc} and filtered by Trimmomatic v.0.39 \citep{bolger2014trimmomatic}. 
      Then clean reads are mapped to host genome using Bowtie2 v.2.4.2 \citep{langmead2012fast} and non-host reads are extracted by SAMtools v.1.11 \citep{li2009sequence}. 
      \newline
      The non-host reads are subject to the module of assembly-dependent species identification. 
      \textit{De novo} assembly is conducted by SPAdes v.3.15.2 \citep{prjibelski2020using}. 
      Assembled contigs are aligned to NCBI non-redundant (nr) database by DIAMOND v.2.0.7.145 \citep{buchfink2015fast}, and assigned to taxa by MEGAN6 \citep{huson2007megan}. 
      \newline
      Then fragment recruitment is conducted to filter reads mapped to species represented by the assembly. 
      A reference database is constructed and it comprises reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding genome annotation in general feature format (gff). 
      For each species represented by assembly, its reference genome dataset, if available, is downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
      Then non-host reads are mapped to the reference database by Bowtie2, and unmapped reads are extracted by SAMtools. 
      \newline
      Reads not recruited by the reference database are subjected to assembly-independent species identification. 
      They are aligned to NCBI nr database through DIAMOND and assigned to taxa by MEGAN6.
      \newline
      Finally, functional annotation is conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      It takes coding sequences (CDSs) of genomes in the reference database and reads subject to assembly-independent species identification as input and assigns them to Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}.
      \newline
      The integrated pipeline was used for analysing metagenomic datasets involved in this study. 
      Details in the pipeline and parameter settings of each module are described in Supplementary \ref{ParameterSetting}.

      \subsubsection{Taxon/function quantification and metabolic pathway reconstruction}
      After profiling, identified species and KOs were quantified by calculating relative sequence abundance, \textit{i.e.} proportion of reads assigned to a species/KO in all reads annotated. 
      For species without available reference genomes, their abundances were calculated using reads assigned to them in assembly-independent search. 
      As for taxa with available reference genomes, they may be identified in both assembly-dependent and -independent search due to strain-specific genomic structures that are not present in reference genomes. 
      Their abundances were calculated by summating number of reads that (1) mapped to coding sequences (CDSs) of reference genomes and (2) assigned to them in assembly-independent search. 
      Reads mapped to non-coding regions were not taken into consideration in order to avoid overestimation since the assembly-independent search was based on aligning reads to nr database, which is composed of proteins. 
      As for KO quantification, CDSs with zero coverage were excluded. 
      Abundances of KOs were calculated by summating number of reads that (1) mapped to CDSs assigned to KOs and (2) assigned to KOs directly. 
      Extraction of CDSs and calculation of their coverage were conducted by BEDtools v.2.30.0 \citep{quinlan2010bedtools}.
      % \newline
      % In order to investigate metabolic potentialities of metagenome, CDSs with non-zero coverage from reference genomes and reads subject to assembly-independent search were functionally annotated using EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      % It searches query sequences  
      % Identified functional categories were quantified by proportion of reads assigned to a category in all reads annotated (relative sequence abundance). 
      % The number of reads assigned to a functional category equals the sum of reads (1) assigned to the category directly and (2) mapped to CDSs assigned to the category.
      \newline
      Metabolic pathways were inferred based on KOs. 
      Reads assigned to plants and arthropods were not included since they were unlikely to represent living organisms. 
      MinPath v.1.6 was used for pathway inference \citep{ye2009parsimony}. 
      It finds a minimal set of KEGG pathways that can explain all KOs provided as input.

@@

      \subsubsection{Simulating different sequencing depth by subsampling and diversity measured using Hill numbers}
      Species/KO inventories obtained from different sequencing depth was simulated by rarefaction. 
      % \textit{i.e.} randomly taking a proportion of the original dataset without replacement. 
      Since the ratio between numbers of raw and clean reads is dependent on sequencing process and is not influenced by sample type, sequencing depth here refers to number of clean read pairs to exclude variance caused by different proportion of low quality reads in samples of same type. 
      Besides, the expected ratio between host and non-host reads in a metagenomic dataset is dependent on the DNA sample and not impacted significantly by sequencing depth. 
      Thus the proportion of non-host reads in each simulation is expected to be the same with that in the original dataset. 
      \newline
      Based on these considerations, I randomly subsampled non-host dataset of each sample, taking 10\%-100\% of read pairs at interval of 10\% by \textit{reformat.sh} script of BBmap v.38.90 \citep{bushnell2014bbmap}, and profiled subsampled datasets by the integrated pipeline (Figure \ref{Pipeline}). 
      The sequencing depth of each subsampled dataset equals number of subsampled non-host read pairs divided by ratio between non-host and clean read pairs. 
      Thus, each subsampled dataset of non-host reads is corresponded to an imaginary dataset of clean reads, whose proportion of non-host reads is the same with that of the original metagenomic dataset. 
      \newline
      After profiling subsampled datasets, species/KO diversity was measured by Hill number, a parametric family of diversity indexes differing by order $q$ \citep{hill1973diversity,ma2018measuring,roswell2021conceptual}. 
      Hill numbers provide unique advantages over other diversity indexes \citep{chao2014unifying,chao2014rarefaction,alberdi2019guide,roswell2021conceptual}. 
      First, Hill numbers follow a replication principle: if a proportion of categories in an inventory was removed randomly, all Hill numbers are expected to decrease by that proportion. 
      Second, the sensitivity of Hill numbers to relative abundances can be modulated by order $q$. 
      % Third, all Hill numbers are pressed in the units of effective numbers of species: the number of equally abundant species in an imaginary inventory with the same diversity. 
      Third, widely used diversity indexes including species richness, Shannon index and Simpson index, can be converted to Hill numbers through algebraic transformations. 
      \newline
      Hill number of order $q$ is defined as Equation \ref{HillNumbers} \citep{hill1973diversity}.
      \begin{equation}
        D^{(q)} = (\sum_{i}(p_i)^q)^{\frac{1}{1-q}}
        \label{HillNumbers}
      \end{equation}
      $p_i$ represents the relative abundance of $i$th species/KO. 
      When $q = 0$, abundances are not taken into consideration and $D^{(0)}$ equals species/KO richness. 
      When $q = 1$, emphasis is given to species/KOs with general abundances. 
      Hill number is defined as the limit of Equation \ref{HillNumbers} as $q$ tends to 1 and equals the exponential of Shannon index (\ref{HillShannon}).  
      \begin{equation}
        D^{(1)} = e^{-\sum_{i}p_ilogp_i}
        \label{HillShannon}
      \end{equation}
      When $q = 2$, high leverage is provided to abundant species/KOs and Hill number equals the inverse of Simpson index (Equation \ref{HillSimpson}).
      \begin{equation}
        D^{(2)} = \frac{1}{\sum_{i}(p_i)^2}
        \label{HillSimpson}
      \end{equation}

@@

      All Hill numbers behave in the following way: removing a proportion of categories in an inventory randomly is expected to cause decrease of Hill numbers in that proportion \citep{roswell2021conceptual}. 

@@

    Fungal and virus species were identified. 

@@

    analysing taxonomic and functional diversity, rarefaction analysis was conducted. 

@@

      \caption{Rarefaction curves for species (a, b, c) or Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) (d, e, f) richness (Hill number of order 0). 
      The horizontal axis represents sequencing depth, and the vertical axis represents richness. 

@@

      Note that the scale of horizontal axis in subfigure c and f is much smaller than that in other subfigures.}

@@

      \textit{Bee\_Bimpatiens\_hv3\_1}&2&8.73   &2.20 &6.16 &0.7051  &6.28 &7.06 &0.8088  &9.45 &7.28 &0.8344  &23.89&7.60 &0.8708\\
      \textit{Bee\_Bimpatiens\_hv4\_1}&2&8.33   &3.12 &6.28 &0.7540  &8.52 &7.50 &0.9002  &12.30&7.76 &0.9323  &26.68&8.09 &0.9718\\
      \textit{Bee\_Bimpatiens\_hv4\_2}&2&7.12   &2.90 &5.34 &0.7493  &7.47 &6.38 &0.8961  &10.58&6.60 &0.9270  &23.73&6.90 &0.9683\\          
      \hline
      \end{tabular}%
      }
      \end{table}
